---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# isofracDH

<!-- badges: start -->
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![R](https://img.shields.io/badge/language-R-blue)]()
<!-- badges: end -->

### Tidyverse tools for hydrogen isotope fractionation calculations

isofracDH provides tidyverse-native functions for calculating hydrogen isotope fractionation and mixing relationships described in:

> Sessions, A. L. (2005). Calculation of hydrogen isotopic fractionations in biogeochemical systems. Geochimica et Cosmochimica Acta.

This package is designed as a teaching and analysis tool for students and researchers learning about hydrogen isotope systematics, with a focus on Sections 3.1, 3.2, 3.3, and 3.5 of Sessions and Hays (2005).

## Features
isofracDH helps you:

- calculate substrate → product fractionation relationships
- model two-component mixing where one component fractionates
- model two-component mixing where both components fractionate
- fit the p-parameter regression formulation described in the paper
- visualize solution constraints and non-unique solution spaces
- reproduce the *Methylococcus capsulatus* analysis from Sessions (2002) 

The package includes:

- tidyverse-friendly calculation functions
- plots to help visualize hydrogen isotope systematics
- simulated teaching datasets (e.g., kerogen exchangeable H examples)
- a packaged dataset derived from Table 3 (*M. capsulatus* lipid p-parameters)


## Installation

You can install the development version of isofracDH from [GitHub](https://github.com/) with:
```{r, eval = F}
# install.packages("pak")
pak::pak("carolynnharris/isofracDH")
```


## Example

This is a basic example showing the intended tidyverse-native workflow (functions add columns to a tibble).

```{r, eval = F}
library(isofracDH)
library(dplyr)

df <- tibble::tibble(
  delta_source = seq(-200, 50, by = 10),
  alpha = 0.92
)

df %>%
  add_frac_single(delta_source = delta_source, alpha = alpha)
```


## Important notes on units 

All delta (δ) values should be supplied in per mille (‰) units.

Internally, isofracDH converts δ values to fractional form (δ / 1000) before applying subsequent equations. This matches the convention used in Sessions and Hays 2005 and helps avoid common unit mistakes.

## Data
The package includes a dataset derived from Table 3 in Sessions et al. (2002) as reproduced in Sessions and Hays (2005), used for the *Methylococcus capsulatus* lipid example in Section 3.5.

## Status

This package is under active development. Planned milestones:

- Implement calculations and plots corresponding to Sections 3.1, 3.2, 3.3, and 3.5
- Add teaching vignettes (single-component fractionation; exchangeable H in kerogen; non-unique solutions; *M. capsulatus* example)
- Add test coverage with testthat

## License
MIT License © Carolynn Harris

## Citation
If you use this package, please cite:

> Harris, C. (2026). isofracDH: Tidyverse tools for hydrogen isotope fractionation calculations. GitHub repository.
